A Framework for Substructural Type Systems \thanks{Supported by organization x.}

### 2021-10-08 tex/esop22/syntax.tex
[ new ] switch to working on ESOP submission
[ writing ] add example term

\begin{figure}
  \begin{align*}
    \begin{prooftree}
      \hypo{\grR \leq \bra x}
      \hypo{\gamma_x = A}
      \infer2[Var]{\grR\gamma \vdash A}
    \end{prooftree}
  \end{align*}
  \begin{align*}
    \begin{prooftree}
      \hypo{\grR\gamma, \gr1A \vdash B}
      \infer1[$\multimap$-I]{\grR\gamma \vdash A \multimap B}
    \end{prooftree}
    &&
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \hypo{\grP\gamma &\vdash A \multimap B}
      \infer[no rule]1{\grQ\gamma &\vdash A \phantom{{} \multimap B}}
      \infer2[$\multimap$-E]{\grR\gamma \vdash B}
    \end{prooftree}
  \end{align*}
  \begin{align*}
    \begin{prooftree}
      \hypo{\grR\gamma \vdash A_i}
      \infer1[$\oplus$-I$_i$]{\grR\gamma \vdash A_0 \oplus A_1}
    \end{prooftree}
    &&
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \infer[no rule]1{\grP\gamma \vdash A \oplus B}
      \hypo{\grQ\gamma, \gr1A &\vdash C}
      \infer[no rule]1{\grQ\gamma, \gr1B &\vdash C}
      \infer2[$\oplus$-E]{\grR\gamma \vdash C}
    \end{prooftree}
  \end{align*}
  \begin{align*}
    \begin{prooftree}
      \hypo{\grR \leq \gr r\grP}
      \hypo{\grP\gamma \vdash A}
      \infer2[$\oc_{\gr r}$-I]{\grR\gamma \vdash \oc_{\gr r} A}
    \end{prooftree}
    &&
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \hypo{\grP\gamma \vdash \oc_{\gr r} A}
      \infer[no rule]1{\grQ\gamma, \gr rA \vdash C}
      \infer2[$\oc_{\gr r}$-E]{\grR\gamma \vdash C}
    \end{prooftree}
  \end{align*}
  \caption{A fragment of \name{}}\label{fig:lr}
\end{figure}

\begin{figure}
  \begin{align*}
    \begin{prooftree}[comb]
      \hypo{\sqni A}
      \infer1[Var]{\vdash A}
    \end{prooftree}
    &&
    \begin{prooftree}[comb]
      \hypo{\gr1A \vdash B}
      \infer1[$\multimap$-I]{\vdash A \multimap B}
    \end{prooftree}
    &&
    \begin{prooftree}[comb]
      \hypo{\vdash A \multimap B}
      \hypo{\sep}
      \hypo{\vdash A}
      \infer3[$\multimap$-E]{\vdash B}
    \end{prooftree}
  \end{align*}
  \begin{align*}
    \begin{prooftree}[comb]
      \hypo{\vdash A_i}
      \infer1[$\oplus$-I$_i$]{\vdash A_0 \oplus A_1}
    \end{prooftree}
    &&
    \begin{prooftree}[comb]
      \hypo{\vdash A \oplus B}
      \hypo{\sep}
      \hypo{(\gr1A \vdash C}
      \hypo{\dottimes}
      \hypo{\gr1B \vdash C)}
      \infer5[$\oplus$-E]{\vdash C}
    \end{prooftree}
  \end{align*}
  \begin{align*}
    \begin{prooftree}[comb]
      \hypo{\gr r \cdot \plr{\vdash A}}
      \infer1[$\oc_{\gr r}$-I]{\vdash \oc_{\gr r} A}
    \end{prooftree}
    &&
    \begin{prooftree}[comb]
      \hypo{\vdash \oc_{\gr r} A}
      \hypo{\sep}
      \hypo{\gr rA \vdash C}
      \infer3[$\oc_{\gr r}$-E]{\vdash C}
    \end{prooftree}
  \end{align*}
  \caption{\cref{fig:lr} restated in terms of bunched combinators}
  \label{fig:lr-comb}
\end{figure}

@@

\missingfigure{Example term}

>>>>

\begin{figure}
  \begin{displaymath}
    \begin{prooftree}
      \hypo{\grR \leq \bra x}
      \hypo{\gamma_x = A}
      \infer2[Var]{\grR\gamma \vdash A}
    \end{prooftree}
  \end{displaymath}
  \begin{displaymath}
    \begin{prooftree}
      \hypo{\grR\gamma, \gr1A \vdash B}
      \infer1[$\multimap$-I]{\grR\gamma \vdash A \multimap B}
    \end{prooftree}
    \qquad
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \hypo{\grP\gamma &\vdash A \multimap B}
      \infer[no rule]1{\grQ\gamma &\vdash A \phantom{{} \multimap B}}
      \infer2[$\multimap$-E]{\grR\gamma \vdash B}
    \end{prooftree}
  \end{displaymath}
  \begin{displaymath}
    \begin{prooftree}
      \hypo{\grR\gamma \vdash A_i}
      \infer1[$\oplus$-I$_i$]{\grR\gamma \vdash A_0 \oplus A_1}
    \end{prooftree}
    \qquad
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \infer[no rule]1{\grP\gamma \vdash A \oplus B}
      \hypo{\grQ\gamma, \gr1A &\vdash C}
      \infer[no rule]1{\grQ\gamma, \gr1B &\vdash C}
      \infer2[$\oplus$-E]{\grR\gamma \vdash C}
    \end{prooftree}
  \end{displaymath}
  \begin{displaymath}
    \begin{prooftree}
      \hypo{\grR \leq \gr r\grP}
      \hypo{\grP\gamma \vdash A}
      \infer2[$\oc_{\gr r}$-I]{\grR\gamma \vdash \oc_{\gr r} A}
    \end{prooftree}
    \qquad
    \begin{prooftree}
      \hypo{\grR \leq \grP + \grQ}
      \hypo{\grP\gamma \vdash \oc_{\gr r} A}
      \infer[no rule]1{\grQ\gamma, \gr rA \vdash C}
      \infer2[$\oc_{\gr r}$-E]{\grR\gamma \vdash C}
    \end{prooftree}
  \end{displaymath}
  \caption{A fragment of \name{}}\label{fig:lr}
\end{figure}

\begin{figure}
  \begin{displaymath}
    \begin{prooftree}[comb]
      \hypo{\sqni A}
      \infer1[Var]{\vdash A}
    \end{prooftree}
    \qquad
    \begin{prooftree}[comb]
      \hypo{\gr1A \vdash B}
      \infer1[$\multimap$-I]{\vdash A \multimap B}
    \end{prooftree}
    \qquad
    \begin{prooftree}[comb]
      \hypo{\vdash A \multimap B}
      \hypo{\sep}
      \hypo{\vdash A}
      \infer3[$\multimap$-E]{\vdash B}
    \end{prooftree}
  \end{displaymath}
  \begin{displaymath}
    \begin{prooftree}[comb]
      \hypo{\vdash A_i}
      \infer1[$\oplus$-I$_i$]{\vdash A_0 \oplus A_1}
    \end{prooftree}
    \qquad
    \begin{prooftree}[comb]
      \hypo{\vdash A \oplus B}
      \hypo{\sep}
      \hypo{(\gr1A \vdash C}
      \hypo{\dottimes}
      \hypo{\gr1B \vdash C)}
      \infer5[$\oplus$-E]{\vdash C}
    \end{prooftree}
  \end{displaymath}
  \begin{displaymath}
    \begin{prooftree}[comb]
      \hypo{\gr r \cdot \plr{\vdash A}}
      \infer1[$\oc_{\gr r}$-I]{\vdash \oc_{\gr r} A}
    \end{prooftree}
    \qquad
    \begin{prooftree}[comb]
      \hypo{\vdash \oc_{\gr r} A}
      \hypo{\sep}
      \hypo{\gr rA \vdash C}
      \infer3[$\oc_{\gr r}$-E]{\vdash C}
    \end{prooftree}
  \end{displaymath}
  \caption{\cref{fig:lr} restated in terms of bunched combinators}
  \label{fig:lr-comb}
\end{figure}

@@

Here is an example term, using the \AgdaFunction{$\lambda$R} system.
First, for ease of writing, we introduce pattern synonyms for each of the
typing rules we use.

\ExecuteMetaData[\PaperExamplestex]{patterns}

Our example term is a function that flips a tagged union wrapped in an
arbitrarily annotated \emph{bang}.
Much of the effort in writing such a term goes into writing the various
relatedness proofs between usage contexts --- observing, for example, that two
usage contexts sum together to make a third, or that a usage context used for
a variable is a basis vector.
We give a method of automating these proofs in \cref{sec:usage-elaborator}.

\ExecuteMetaData[\PaperExamplestex]{lR-term}
